---
title: "`r sprintf('Retail Transactions: Monthly Report for Month %s', params$report_month)`"
author: "Christian Wiloejo"
output: html_document
params: 
  report_month: 1
  views_file_path: 'views -- January 2020.csv'
  transactions_file_path: 'transactions -- January 2020.csv'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r seed}
set.seed(41)
```

```{r libraries}
library(data.table)
library(DT)
library(Hmisc)
library(rmarkdown)
```

```{r constants}
# These constants need to be changed for automatic reporting in another R file
report_month = params$report_month
main_path = '../data and templates/'
views_file_path = params$views_file_path
transactions_file_path = params$transactions_file_path
```

```{r functions}
round.numerics <- function(x, digits){
  if(is.numeric(x)){
    x <- round(x = x, digits = digits)
  }
  return(x)
}

#q1
first_week_most_viewed_products_category = function(dat, topN){
  # Extract day and filter for first week
  dat[, day:= as.integer(format(time, format = "%d"))]
  dat = dat[day <= 7]
  
  # Get result with group by and count 
  result = dat[,.(count_views = .N),by=.(product_id, category)][order(count_views, decreasing = TRUE)][1:topN]
  
  return(result)
}

#q2
month_most_viewed_products = function(dat, category_selected, topN){
  dat = dat[category == category_selected]
  result = dat[,.(count_views = .N),by=.(product_id)][order(count_views, decreasing = TRUE)][1:topN]
  
  return(result)
}

#q3
month_revenue_by_category = function(dat){
  dat[,revenue := price*quantity]
  result = dat[, .(total_revenue = sum(revenue)), by=category][order(total_revenue, decreasing = T)]
  
  return(result)
}

#q4
summary_month_customer_purchase = function(dat){
  # Customers with one transaction is equal to transaction data
  dat[,month := as.integer(format(time, "%m"))]
  # Calculate total spending per row
  dat[,purchase_total:=price*quantity]
  # Average, median, and standard deviation of the customers' monthly spending
  result = dat[,.(mean_purchase = mean(purchase_total),
                         median_purchase = median(purchase_total),
                         std_purchase = sd(purchase_total)
                         ), by = .(month)]
  return(result)
}

#q5
pct_dist_spending_by_gender = function(transactions_dat){
  result = transactions_dat[, .(revenue = sum(revenue)), by=gender]
  result[, percentage_distribution :=  revenue/sum(revenue)]
  return(result)
}

#q8 
summary_nunique_products_viewed_per_customer = function(dat){
  # Cut to age groups
  dat[,age_group:=cut2(dat$age, cuts = c(18,35,50,65, max(dat$age)+1))]
  # Find unique customer and product; then count the unique products they viweed
  unique_products_viewed = unique(dat[,c("customer_id", "product_id", "age_group")])
  unique_products_viewed = unique_products_viewed[,.(number_unique_products_viewed = .N),by=.(customer_id, age_group)]
  
  # Compute stats summary result by age group
  result = unique_products_viewed[, .(mean_unique_products_viewed = mean(number_unique_products_viewed), 
                             median_unique_products_viewed = median(number_unique_products_viewed), 
                             std_unique_products_viewed = sd(number_unique_products_viewed))
                         ,by = .(age_group)]
  # Sort by age group
  result = result[order(age_group, decreasing = F)]
  
  return(result)
}

#q9
correlation_total_views_total_spending  = function(views_dat, transactions_dat){
  # Compute total views and total spending 
  total_page_views_per_customer = views_dat[, .(total_views = .N), by = customer_id]
  total_spending_per_customer = transactions_dat[,.(total_spending = sum(revenue)), by = customer_id]
  # Left join total spending to total views 
  dat = merge(total_page_views_per_customer, total_spending_per_customer, by = 'customer_id', all.x = T)
  # Spending = 0 for customers who viewed but didn't purchase
  dat[is.na(total_spending), total_spending := 0]
  # Compute correlation
  result = cor(dat$total_views, dat$total_spending)
  
  return(result)
}

#q10
largest_coat_purchase = function(transactions_data){
  just_coats_transactions = transactions_data[category == 'coat', .(product_id, customer_id, category)]
  coats_purchase_per_customer = just_coats_transactions[,.(coats_purchased = .N),by=customer_id]
  max_coats_purchased = max(coats_purchase_per_customer$coats_purchased)
  result = coats_purchase_per_customer[coats_purchased == max_coats_purchased]
  return(result)
}

```

```{r load_data}
# Load raw data
customers_dt = fread(paste0(main_path, 'customers.csv'))
products_dt = fread(paste0(main_path, 'products.csv'))

views_dt = fread(paste0(main_path, views_file_path))
transactions_dt = fread(paste0(main_path, transactions_file_path))
```

```{r explore_data, eval = FALSE}
```


```{r clean_data}
# Data Type Handling
views_dt[, time := as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%S", tz = 'UTC')]
customers_dt[, gender := as.factor(gender)]
customers_dt[, region := as.factor(region)]
products_dt[, category := as.factor(category)]

# Remove following month's data in current month's dataset (example: February data in January data or March data in February dataset)
transactions_dt[, month := as.integer(format(time, '%m'))]
transactions_dt = transactions_dt[month == report_month]

views_dt[, month := as.integer(format(time, '%m'))]
views_dt = views_dt[month == as.name(params$report_month)]

# Joins each views and transactions to customers and products data
views_full_dat = merge(x = views_dt, y = customers_dt, by = 'customer_id', all.x = TRUE)
views_full_dat = merge(x = views_full_dat, y = products_dt, by = 'product_id', all.x = TRUE)
transactions_full_dat = merge(x = transactions_dt, y = customers_dt, by = 'customer_id', all.x = TRUE)
transactions_full_dat = merge(x = transactions_full_dat, y = products_dt, by = 'product_id', all.x = TRUE)

# Check if price data in transactions are equal
if (mean(transactions_full_dat$price.x == transactions_full_dat$price.y) == 1){
  # print("Price data is equal")
  # Removing duplicate columns in transactions data and renaming
  transactions_full_dat[,price.y := NULL]
  setnames(transactions_full_dat, "price.x", "price")

} else {
  print("Price data is not equal. Proceed but double check why that is")
}

```



## Part 2:  Specific Questions {.tabset}

This part of the report will be directed to product managers throughout the client's company.  The idea is to give them the useful information they need to act on the specific questions they posed.  Plan your communication accordingly.

For this report, make sure to use all of the data that was provided to you for the month.  If you do note any issues with the data (Part 3), this can be reported to the engineering team for them to resolve.

### Q1

#### Question

During the first week of the month, what were the 10 most viewed products?  Show the results in a table with the product's identifier, category, and count of the number of views.

#### Answer

```{r part2_q1}
datatable(first_week_most_viewed_products_category(views_full_dat, 10))
```


### Q2

#### Question

During the whole month, what were the 10 most viewed products for each category?  Show the results in separate tables by category.  Including only the product's identifier and the count of the number of views.


#### Answer {.tabset}
##### Shirt

```{r part2_q2_shirt}
datatable(month_most_viewed_products(views_full_dat, "shirt", 10))
```

##### Pants

```{r part2_q2_pants}
datatable(month_most_viewed_products(views_full_dat, "pants", 10))
```

##### Coat

```{r part2_q2_coat}
datatable(month_most_viewed_products(views_full_dat, "coat", 10))
```

##### Shoes

```{r part2_q2_shoes}
datatable(month_most_viewed_products(views_full_dat, "shoes", 10))
```

##### Hat

```{r part2_q2_hat}
datatable(month_most_viewed_products(views_full_dat, "hat", 10))
```


### Q3

#### Question

What was the total revenue for each category of product during the month?  Show the results in a single table sorted in decreasing order.

#### Answer

```{r part2_q3}
datatable(month_revenue_by_category(transactions_full_dat))
```

### Q4

#### Question

Among customers with at least one transaction, show the average, median, and standard deviation of the customers' monthly spending on the site.

#### Answer

```{r part2_q4}
datatable(summary_month_customer_purchase(transactions_full_dat))
```

### Q5

#### Question

What is the percentage distribution of spending by gender?  Show the amount of revenue and the percentage.

#### Answer

```{r part2_q5}
datatable(pct_dist_spending_by_gender(transactions_full_dat))
```

### Q6

#### Question

Using linear regression, what is the effect of an extra ten thousand dollars of income on monthly spending for a customer while adjusting for age, gender, and region?

#### Answer

```{r part2_q6}
linear_model = summary(lm(revenue ~ income+age+gender+region, data = transactions_full_dat))
coefs = linear_model$coefficients
print(paste0("While holding age, gender, and region, every extra ten thousand dollars of income would increase spending for a customer by $", round.numerics(coefs[2,1]*10000,2)))
```


### Q7

#### Question

Among customers who viewed at least 1 product, how many had at least one purchase during the month?  Show the total number and as a percentage of the users with a view.

#### Answer

```{r part2_q7}
numerator = length(unique(views_full_dat[customer_id %in% unique(transactions_full_dat$customer_id)]$customer_id))
denominator = length(unique(views_full_dat$customer_id))

print(paste0("The total number of those who viewed at least 1 product and purchased during the month is ", numerator, " which is ", round.numerics((numerator/denominator)*100,2), "% of the users with a view"))
```

### Q8

#### Question

Now let's look at the viewing habits in different age groups, including 18-34, 35-49, 50-64, and 65+.  Within each group, what were the mean, median, and standard deviation for the number of unique products viewed per customer?

#### Answer

```{r part2_q8}
datatable(summary_nunique_products_viewed_per_customer(views_full_dat))
```


### Q9 

#### Question

What is the correlation between a user's total page views and total spending?  For customers without a transaction, include their spending as zero.

#### Answer

```{r part2_q9}
datatable(transactions_full_dat)
```
```{r}
print(paste0("The correlation between user's total views and total spending is ", round.numerics(correlation_total_views_total_spending(views_full_dat, transactions_full_dat),2)))
```

### Q10

#### Question

Which customer purchased the largest number of coats?  In the event of a tie, include all of the users who reached this value.  Show their identifiers and total volume.

#### Answer

```{r part2_q10}
datatable(largest_coat_purchase(transactions_full_dat))
```

